---
title: "Geomapping"
author: "Alexander Ng"
date: "10/14/2021"
output: 
  html_document:
    df_print: paged
    highlight: tango
    number_sections: yes
    theme: readable
    toc: yes
    toc_depth: 2
    toc_float: no
    fontsize: 12

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(kableExtra)
library(sf)
library(ggspatial)

```
# Loading the Map Data Files

```{r define-dirs}
proj_dir = "/Volumes/GDRIVE_SSD/homes/alex/datascience/698_CAPSTONE_2021_FALL"

working_dir = paste0(proj_dir, "/", "project_code/Part2_GeoMapping/")

data_dir = paste0(proj_dir, "/project_data/")

setwd(working_dir)

```

```{r map-file-load}
raleigh_corporate_limits_file = paste0(data_dir, "Corporate_Limits.geojson")

cityworks_potholes_file = paste0(data_dir, "Cityworks_Potholes.geojson" )

raleigh_parks_file = paste0(data_dir, "Raleigh_Parks.geojson")
```

```{r load-potholes}
gj_potholes = sf::st_read(cityworks_potholes_file, quiet = TRUE ) %>% st_transform(2264)

```


```{r load-parks}
gj_parks = sf::st_read(raleigh_parks_file, quiet = TRUE ) %>% st_transform(2264)

```


```{r load-raleigh-boundary}
#
# Load the geojson files
# change the coordinate system to 2264
# filter objects that belong to RALEIGHT
# --------------------------------------------
gj_raleigh <- sf::st_read(raleigh_corporate_limits_file, quiet = TRUE) %>% 
  st_transform(2264) %>% 
  filter( SHORT_NAME == 'RAL' )

head(gj_raleigh) %>% select(-ORDINANCE_NUMBER, -EFFECTIVE_DATE, -SHAPEAREA ) %>% st_drop_geometry() %>%
  kable() %>% kable_styling(bootstrap_options = c("striped", "hover"))
```

We compute the area of the polygonal representation and validate by checking against Wikipedia's reported area.
They are very close.

```{r}
# Compute the total square miles of the polygons.
# The NAD83 North Carolina projection uses square ft.
# ------------------------------------------------------
area_of_polygons = st_area(gj_raleigh)

sprintf( "Raleigh Area is: %.2f square miles.  Wikipedia gives: 147.64 sq mi" , sum(area_of_polygons) / ( 5280^2)  )

```

The square mile area for most areas look accurate.  However, note that they show area inside
Wake County.  Thus, Durham has less than 1 mile inside Wake County but actual area is over 100 squared miles.

```{r}
a = sf::st_read(raleigh_corporate_limits_file, quiet = TRUE) %>%   st_transform(2264) 

a = a %>% mutate( my_area = st_area(a)) %>% mutate( sqmiles = my_area/ (5280^2)) %>% st_drop_geometry() %>%
   select(SHORT_NAME, LONG_NAME, sqmiles) %>% group_by(SHORT_NAME, LONG_NAME) %>% summarize( sqmiles = sum( sqmiles)) 

attributes( a$sqmiles ) = NULL

a %>% kable(digits = 2, caption = "Square Miles of Towns inside Wake County")

```

# Visualization with GGPLOT and related libraries

## Plotting the City and Base Layer

ggplot and geom_sf with no other arguments.

```{r}
gj_raleigh %>% ggplot() + geom_sf() + labs(title = "Raleigh - Plain Map")
```


Fill the interior with color and add a border in red.

```{r}
gj_raleigh %>% ggplot() + geom_sf( fill = "skyblue", color = "darkred", size = 0.1) + 
  labs(title = "Raleigh City Limits in Color")
```

Let's put a basemap tile underneath and  compass rose and alpha blend.

```{r}
gj_raleigh %>% ggplot() + geom_sf( fill = "skyblue", color = "darkred") + 
  labs(title = "Raleigh City:  Basemap , Compass, Alpha Blending") +
  annotation_map_tile(type="osm", progress = "none", alpha = 0.5 ) +
  annotation_north_arrow(location="bl", which_north = "true")
```

## Plotting Point Data and Cities As Layers

Here is a plot of all potholes in Raleigh Cityworks database.

```{r}
gj_raleigh %>% ggplot() + geom_sf( fill = "skyblue", color = "darkred") + 
  labs(title = "Raleigh City: Basemap, Boundary and Potholes") +
  geom_sf(data=gj_potholes, color = "yellow" , size = 0.1 ) +
  annotation_map_tile(type="osm", progress = "none", alpha = 0.5 )
```

## Mapping Parks and Cities as Layers


Here is a plot of all city parks in Raleigh database.

```{r out.width="100%"}
gj_raleigh %>% ggplot() + geom_sf( fill = "skyblue", color = "darkred") + 
  labs(title = "Raleigh City: Basemap, Boundary and Parks") +
  geom_sf(data=gj_parks, color = "darkgreen" , fill = "darkgreen") +
  annotation_map_tile(type="osm", progress = "none", alpha = 0.5 )
```

It appears that Forest Ridge Park (in North Raleigh) and Lake Wheeler Park (South Raleigh) are outside of 
City limits.   Lake Wheeler is a large reservoir (650 acres water and 150 acres park).  It is the primary water supply for Raleigh.  It may be desirable to drop these two large parks since they fall outside the town limits.


```{r}
gj_parks %>% select( NAME, PARK_TYPE, CAC, MAP_ACRES) %>% 
  st_drop_geometry() %>% arrange(desc(MAP_ACRES)) %>%
  kable(caption="Parks in Raleigh by Area", digits = 1 ) %>%
  kable_styling(bootstrap_options = c("striped", "hover"))

```

## Geomapping of Assault Incidents in Raleigh


```{r}

gj_assault_geocoded = st_read(paste0(data_dir, "EXP_EDA_RALEIGH_assaults_Aug2021.geojson" ) )

head(gj_assault_geocoded)
```

```{r}


gj_raleigh %>% ggplot() + geom_sf( fill = "skyblue", color = "darkred") + 
  labs(title = "Raleigh City: Basemap, Boundary and Crime") +
  geom_sf(data=gj_assault_geocoded, color = "yellow" , size = 0.1 ) +
  annotation_map_tile(type="osm", progress = "none", alpha = 0.5 )
```

## Crime Density Maps by Year


### Crime in 2014

This was a partial year
```{r}

gj_raleigh %>% ggplot() + geom_sf( fill = "skyblue", color = "darkred") + 
  labs(title = "Raleigh City: Crime Map 2014") +
  layer_spatial(data=gj_assault_geocoded %>% filter(reported_year==2014), color = "yellow" , size = 0.1 ) +
  annotation_map_tile(type="osm", progress = "none", alpha = 0.5 ) 


```


### Crime in 2015

This was a full year
```{r}
this_year = 2015

gj_raleigh %>% ggplot() + geom_sf( fill = "skyblue", color = "darkred") + 
  labs(title = paste0("Raleigh City: Crime Map ", this_year ) ) +
  layer_spatial(data=gj_assault_geocoded %>% filter(reported_year==this_year), color = "yellow" , size = 0.1 ) +
  annotation_map_tile(type="osm", progress = "none", alpha = 0.5 ) 

```



### Crime in 2016

This was a full year
```{r}
this_year = 2016

gj_raleigh %>% ggplot() + geom_sf( fill = "skyblue", color = "darkred") + 
  labs(title = paste0("Raleigh City: Crime Map ", this_year ) ) +
  layer_spatial(data=gj_assault_geocoded %>% filter(reported_year==this_year), color = "yellow" , size = 0.1 ) +
  annotation_map_tile(type="osm", progress = "none", alpha = 0.5 ) 

```


### Crime in 2017

This was a full year
```{r}
this_year = 2017

gj_raleigh %>% ggplot() + geom_sf( fill = "skyblue", color = "darkred") + 
  labs(title = paste0("Raleigh City: Crime Map ", this_year ) ) +
  layer_spatial(data=gj_assault_geocoded %>% filter(reported_year==this_year), color = "yellow" , size = 0.1 ) +
  annotation_map_tile(type="osm", progress = "none", alpha = 0.5 ) 

```


### Crime in 2018

This was a full year
```{r}
this_year = 2018

gj_raleigh %>% ggplot() + geom_sf( fill = "skyblue", color = "darkred") + 
  labs(title = paste0("Raleigh City: Crime Map ", this_year ) ) +
  layer_spatial(data=gj_assault_geocoded %>% filter(reported_year==this_year), color = "yellow" , size = 0.1 ) +
  annotation_map_tile(type="osm", progress = "none", alpha = 0.5 ) 

```


### Crime in 2019

This was a full year
```{r}
this_year = 2019

gj_raleigh %>% ggplot() + geom_sf( fill = "skyblue", color = "darkred") + 
  labs(title = paste0("Raleigh City: Crime Map ", this_year ) ) +
  layer_spatial(data=gj_assault_geocoded %>% filter(reported_year==this_year), color = "yellow" , size = 0.1 ) +
  annotation_map_tile(type="osm", progress = "none", alpha = 0.5 ) 

```


### Crime in 2020

This was a full year
```{r}
this_year = 2020

gj_raleigh %>% ggplot() + geom_sf( fill = "skyblue", color = "darkred") + 
  labs(title = paste0("Raleigh City: Crime Map ", this_year ) ) +
  layer_spatial(data=gj_assault_geocoded %>% filter(reported_year==this_year), color = "yellow" , size = 0.1 ) +
  annotation_map_tile(type="osm", progress = "none", alpha = 0.5 ) 

```



### Crime in 2021

This was a partial year
```{r}
this_year = 2021

gj_raleigh %>% ggplot() + geom_sf( fill = "skyblue", color = "darkred") + 
  labs(title = paste0("Raleigh City: Crime Map ", this_year ) ) +
  layer_spatial(data=gj_assault_geocoded %>% filter(reported_year==this_year), color = "yellow" , size = 0.1 ) +
  annotation_map_tile(type="osm", progress = "none", alpha = 0.5 ) 

```

## Kernel Density Plots

The KDE code below does not work as modified.  Need to think of a better solution.

```{r eval=FALSE}
ggplot(data=gj_assault_geocoded %>% filter(reported_year==2019) ) +
   stat_sf_coordinates() +
    coord_equal() +
   xlab('Longitude') + ylab('Latitude') +
  stat_density2d(aes(fill = ..level.., x = after_stat(x), y = after_stat(y)), 
                 alpha = 0.5 , h = 0.007, n = 500, geom = "polygon" ) +
  scale_fill_viridis_c() +
  annotation_map_tile(type="osm", progress = "none", alpha = 0.5 ) 
  theme(legend.position = 'none' )


```

