---
title: "EDA Predictions"
author: "Alexander Ng"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document:
    df_print: paged
    highlight: tango
    number_sections: yes
    theme: readable
    toc: yes
    toc_depth: 2
    toc_float: no
    fontsize: 12
  pdf_document:
    toc: yes
    toc_depth: '2'
subtitle: Production
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r message=FALSE, warning=FALSE}
suppressPackageStartupMessages(library(knitr) )
suppressPackageStartupMessages( library(tidyverse) )
suppressPackageStartupMessages(library(kableExtra) )

suppressPackageStartupMessages(library(sf))
suppressPackageStartupMessages(library(caret))
library(ggspatial)
library(cowplot)
library(Cubist)
library(ranger)
library(cowplot)
library(skimr)
library(viridis)
library(RColorBrewer)
library(leaflet)
library(htmltools)
library(htmlwidgets)
library(lubridate)
library(tidygeocoder)

```

# Introduction

The objective of this analysis is to build interactive visualizations of predicted crime hotspots, actual crime behaviour and the predictors in Raleigh, North Carolina.
Due to the large volume of data, we divide the analyses into two maps.  One map depicts the hotspots together with demographic predictors.   The second map depicts the hotspots
together with points of interest predictors and all predicted crime in the city.


# Configuration and File Import

The model output was pre-generated.
We will load the Models for the 2020 test period.   Base date is 2019-12-31.


```{r configuration, echo = FALSE}

proj_dir = "/Volumes/GDRIVE_SSD/homes/alex/datascience/698_CAPSTONE_2021_FALL"
working_dir = paste0(proj_dir, "/", "project_code/Part4_Interactive/")
data_dir = paste0(proj_dir, "/project_data/")

setwd(working_dir)
```



```{r define-model-result-files-2019}

PAI_cubist_ranked_grid_20191231_file = paste0(data_dir, "EXP_CUBIST_PAI_RANKED_GRID_2019-12-31.geojson")

PEI_cubist_csv_grid_20191231_file = paste(data_dir, "EXP_CUBIST_PEI_RANKED_CSV_2019-12-31.csv")

PAI_rf_ranked_grid_20191231_file = paste0(data_dir, "EXP_RANDOMFOREST_PAI_RANKED_GRID_2019-12-31.geojson")

PEI_rf_csv_grid_20191231_file = paste(data_dir, "EXP_RANDOMFOREST_PEI_RANKED_CSV_2019-12-31.csv")

```

```{r map-file-load}
raleigh_corporate_limits_file = paste0(data_dir, "Corporate_Limits.geojson")

```

```{r load-raleigh-boundary}
#
# Load the geojson files
# change the coordinate system to 2264
# filter objects that belong to RALEIGHT
# --------------------------------------------
gj_raleigh <- sf::st_read(raleigh_corporate_limits_file, quiet = TRUE) %>% 
  st_transform(2264) %>% 
  filter( SHORT_NAME == 'RAL' )

gj_raleigh_buffer = st_buffer(gj_raleigh, dist = 100 ) %>% st_union() %>% st_sf() 

```

```{r load-PAI-cubist-grid}

PAI_cu_ranked_grid_20191231 <- sf::st_read( PAI_cubist_ranked_grid_20191231_file, quiet = TRUE ) %>%
    st_transform(2264) 

PAI_rf_ranked_grid_20191231 <- sf::st_read( PAI_rf_ranked_grid_20191231_file, quiet = TRUE ) %>%
    st_transform(2264) 

```

# Properties of Assaults on the Grid

```{r}
PAI_rf_ranked_grid_20191231 %>% st_drop_geometry() %>% 
  group_by( crime_freq ) %>% 
  summarize( count = n() ) %>% 
  mutate( pct_crime = 100 * count/sum(count) ) %>% ggplot(aes( x = reorder( as.factor(crime_freq), desc(crime_freq) ), 
                                                               y = pct_crime )) + geom_bar(stat = "identity") + coord_flip() +
   geom_text(aes(label = round(pct_crime, 2)), hjust = 1 , color = "red") +
  labs(title = "Assaults on Raleigh Grid", subtitle = "Actual 2020 Assaults", x = "Incidents per Grid Cell", y = "% of Total" ) -> p

ranked_grid_assaults_file = paste0(data_dir, "EXP_RANDOMFOREST_BARPLOT_RANKED_GRID_2020.png")

cowplot::save_plot(  ranked_grid_assaults_file, p , base_width = 6  )

p

```

# Sensitivity

```{r}

PAI_rf_ranked_grid_20191231 %>% 
  st_drop_geometry() %>% 
  group_by(hotspot) %>% 
  skim() %>% 
  yank("numeric") %>%
  filter( !( skim_variable %in% c( 'id_grid', 'cols', 'rows', 'rank', 'area', 'lon', 'lat',
                                   'PAI', 'RRI', 'cum_area', 'prop_cum_area' ,
                                   'cum_crime_freq', 'prop_cum_crime_freq') ) ) %>%
  select( skim_variable, hotspot, mean, p50) %>%
  rename( variable = skim_variable) %>% 
  select( variable, hotspot, mean) %>%
  group_by( variable) %>%
  mutate( p = ifelse( hotspot > 0 , mean, (sum(mean)-mean))/sum(mean) ) %>%
  ungroup() 
# %>% filter(variable == 'income')
# %>% pivot_wider( names_from = hotspot, values_from = mean) %>% kable(digits = 3) %>% kable_styling(bootstrap_options = c("hover", "striped"))


```


```{r}
PAI_rf_ranked_grid_20191231 %>% 
  st_drop_geometry() %>% 
  group_by(hotspot) %>% 
  skim() %>% 
  yank("numeric") %>%
  filter( !( skim_variable %in% c( 'id_grid', 'cols', 'rows', 'rank', 'area', 'lon', 'lat', 'crime_count', 
                                   'crime_count_3Y', 'crime_count_2Y', 'crime_count_1Y', 'crime_count_6M', 
                                   'PAI', 'RRI', 'cum_area', 'prop_cum_area' ,
                                   'cum_crime_freq', 'prop_cum_crime_freq') ) ) %>%
  select( skim_variable, hotspot, mean, p50) %>%
  rename( variable = skim_variable) %>% 
  select( variable, hotspot, mean) %>%
  group_by( variable) %>%
  mutate( p = ifelse( hotspot==1 , mean, sum(mean) - mean ) /sum(mean) ) %>%
  ungroup() %>%
  ggplot(aes(fill = as.factor(hotspot), y = mean, x = reorder(variable, p) ) ) + geom_bar(position = "fill", stat="identity", width = 0.8) +
   scale_fill_manual("legend", values = c("0" = "lightblue", "1"="orange")) +
  labs( y = "Proportion", x = "Variable", fill = "Hot Spot", 
        title = "Ratio of Variable for Hotspots vs. Non-Spots" ,
        subtitle = "Random Forest - 2020 test period") +
  theme(aspect.ratio = 1/3, axis.text.x = element_text(angle = 90, hjust = .95, vjust = 0.2 ) ) -> plot_variables_hotspot

plot_variables_hotspot

cowplot::save_plot( paste0(data_dir, "EXP_RANDOMFOREST_PREDICTORS_HOTSPLOT_PLOT_2020.png" ), plot_variables_hotspot  , base_height = 6 )

```

```{r out.width='100%' }
PAI_rf_ranked_grid_20191231 %>% filter(rank <= 10 ) %>%
  select( rank, lon, lat, id_grid, crime_freq, PAI, RRI, prop_cum_area, prop_cum_crime_freq ) %>%  st_drop_geometry() %>%
  tidygeocoder::reverse_geocode(lat = lat, long = lon, address = "address") %>%  
  select(rank, id_grid, crime_freq, PAI, prop_cum_area, prop_cum_crime_freq,address ) %>% 
  kable(digits = 2, caption = "Top 15 Crime Hotspots") %>%
  kable_styling(bootstrap_options = c("hover", "striped"
                                ))
  

```


Save the rank table to file.

```{r out.width='100%' }
PAI_rf_ranked_grid_20191231 %>% filter(rank <= 10 ) %>%
  select( rank, lon, lat, id_grid, crime_freq, PAI, RRI, prop_cum_area, prop_cum_crime_freq ) %>%  st_drop_geometry() %>%
  tidygeocoder::reverse_geocode(lat = lat, long = lon, address = "address") -> PAI_rf_ranked_revgeocode_top10

PAI_rf_ranked_revgeocode_top10 %>% write_rds(paste0(data_dir, "EXP_PAI_RANKED_REVGEOCODE_RANDOMFOREST_2020_TOP10.rds") )
                                

```

```{r}

gj_assault_geocoded = st_read(paste0(data_dir, "EXP_EDA_RALEIGH_assaults_Aug2021.geojson" ), quiet = TRUE )

```

```{r}
grid_cellsize = 490

# We need the grid cell buffer file name for export and later import
grid_buffer_file = paste0(data_dir, "EXP_GRID_RALEIGH_BUFFER_", grid_cellsize, ".geojson")

```


```{r load-city-buffered-grid, echo = FALSE}
# Now we load the grid from file.

grid_raleigh_buffer = sf::st_read(grid_buffer_file, quiet = TRUE ) %>% st_transform(2264)

# Let's check it loading properly

dim(grid_raleigh_buffer)

```


```{r}
gj_assault2grid <- gj_assault_geocoded %>% st_join(grid_raleigh_buffer) %>% filter( !is.na( id_grid  ))


```

```{r}


```