---
title: "Map of Raleigh NC Assaults"
author: "Alexander Ng"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document:
    df_print: paged
    highlight: tango
    number_sections: yes
    theme: readable
    toc: yes
    toc_depth: 2
    toc_float: no
    fontsize: 12
  pdf_document:
    toc: yes
    toc_depth: '2'
subtitle: Production
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r message=FALSE, warning=FALSE}
suppressPackageStartupMessages(library(knitr) )
suppressPackageStartupMessages( library(tidyverse) )
suppressPackageStartupMessages(library(kableExtra) )

suppressPackageStartupMessages(library(sf))
suppressPackageStartupMessages(library(caret))

library(ggspatial)
library(cowplot)
library(Cubist)
library(ranger)
library(cowplot)
library(skimr)
library(viridis)
library(RColorBrewer)
library(leaflet)
library(htmltools)
library(htmlwidgets)
library(lubridate)
```

# {.tabset .tabset-fade}

## TAB 1

# Introduction

The objective of this analysis is to build interactive visualizations of predicted crime hotspots, actual crime behaviour and the predictors in Raleigh, North Carolina.
Due to the large volume of data, we divide the analyses into two maps.  One map depicts the hotspots together with demographic predictors.   The second map depicts the hotspots
together with points of interest predictors and all predicted crime in the city.


# Configuration and File Import

The model output was pre-generated.
We will load the Models for the 2020 test period.   Base date is 2019-12-31.


```{r configuration, echo = FALSE}

proj_dir = "/Volumes/GDRIVE_SSD/homes/alex/datascience/698_CAPSTONE_2021_FALL"
working_dir = paste0(proj_dir, "/", "project_code/Part4_Interactive/")
data_dir = paste0(proj_dir, "/project_data/")

setwd(working_dir)
```



```{r define-model-result-files-2019}

PAI_cubist_ranked_grid_20191231_file = paste0(data_dir, "EXP_CUBIST_PAI_RANKED_GRID_2019-12-31.geojson")

PEI_cubist_csv_grid_20191231_file = paste(data_dir, "EXP_CUBIST_PEI_RANKED_CSV_2019-12-31.csv")

PAI_rf_ranked_grid_20191231_file = paste0(data_dir, "EXP_RANDOMFOREST_PAI_RANKED_GRID_2019-12-31.geojson")

PEI_rf_csv_grid_20191231_file = paste(data_dir, "EXP_RANDOMFOREST_PEI_RANKED_CSV_2019-12-31.csv")

```

```{r map-file-load}
raleigh_corporate_limits_file = paste0(data_dir, "Corporate_Limits.geojson")

```

```{r load-raleigh-boundary}
#
# Load the geojson files
# change the coordinate system to 2264
# filter objects that belong to RALEIGHT
# --------------------------------------------
gj_raleigh <- sf::st_read(raleigh_corporate_limits_file, quiet = TRUE) %>% 
  st_transform(2264) %>% 
  filter( SHORT_NAME == 'RAL' )

gj_raleigh_buffer = st_buffer(gj_raleigh, dist = 100 ) %>% st_union() %>% st_sf() 

```

```{r load-PAI-cubist-grid}

PAI_cu_ranked_grid_20191231 <- sf::st_read( PAI_cubist_ranked_grid_20191231_file, quiet = TRUE ) %>%
    st_transform(2264) 

PAI_rf_ranked_grid_20191231 <- sf::st_read( PAI_rf_ranked_grid_20191231_file, quiet = TRUE ) %>%
    st_transform(2264) 

```

```{r}

# Diagnostic

#PAI_rf_ranked_grid_20191231 %>% 
#  st_drop_geometry() %>% 
#  select( rank, id_grid, cum_crime_freq,  prop_cum_area, prop_cum_crime_freq , PAI , RRI ) %>% 
#  left_join( PAI_cu_ranked_grid_20191231 %>% 
#               st_drop_geometry() %>% 
#               select( rank, id_grid, cum_crime_freq , prop_cum_area, prop_cum_crime_freq, PAI, RRI ), by = "rank") %>%
#  rename( id_grid_rf              = id_grid.x, 
#          cum_crime_freq_rf       = cum_crime_freq.x ,
#          prop_cum_area_rf        = prop_cum_area.x,
#          prop_cum_crime_freq_rf  = prop_cum_crime_freq.x ,
#          PAI_rf                  = PAI.x ,
#          RRI_rf                  = RRI.x ,
#          id_grid_cu              = id_grid.y ,
#          cum_crime_freq_cu       = cum_crime_freq.y ,
#          prop_cum_area_cu        = prop_cum_area.y ,
#          prop_cum_crime_freq_cu  = prop_cum_crime_freq.y ,
#          PAI_cu                  = PAI.y ,
#          RRI_cu                  = RRI.y
#          ) -> PAI_comparison_20191231 

```


# Leaflet Hotspot and Demographics Map

This section builds a map relating crime hotspots to demographics for each of the Random Forest and Cubist models.  


```{r}
# Make a leaflet title

rr <- tags$div(HTML('<h3>Assault Hotspot Prediction</h3><br>2020 Forecasts<br>with Demographic Variables<br>Random Forest Model<br>Raleigh, NC<br>Alexander Ng' ) )



rc <- tags$div(HTML('<h3>Assault Hotspot Prediction</h3><br>2020 Forecasts<br>with Demographic Variables<br>Cubist Model<br>Raleigh, NC<br>Alexander Ng' ) )

```



## Random Forest Map - Demographics

```{r load-rf-grid-output}

# Load the grid cells from random forest. Add HTML table as label when hovering.
# ---------------------------------------------------------------------------------
hotspot_cells_20191231 <- PAI_rf_ranked_grid_20191231 %>% 
                                    filter( hotspot==1) %>% 
                                    st_transform(4326) %>% 
                                    mutate( Label = paste0( "<br><b>Hotspot :</b>",rank , "</b>", 
                                                   "<br><b>","Assaults 2020 ", "</b>" ,
                                                   "<table>" ,
                                                   "<tr><td>", "Predicted:", "</td><td>"    , round(pred_crime_freq, 1) , "</td></tr>" ,
                                                   "<tr><td>", "       2020:", "</td><td>"  , crime_freq , "</td></tr>" ,
                                                   "<tr><td>", "       2019:", "</td><td>"  , round(0.5 * ( crime_freq_6M + crime_freq_1Y), 0) , "</td></tr>" ,
                                                   "<tr><td>", "       2018:", "</td><td>"  , crime_freq_2Y , "</td></tr>" ,
                                                   "<tr><td>", "       2017:", "</td><td>"  , crime_freq_3Y, "</td></tr>" ,
                                                   "</table>"
                                                            ))

#
# Save the crime grid.  Add HTML table for label when hovering over each cell.
# ---------------------------------------------------------------------------------
demographics_cells_20191231 <-  PAI_rf_ranked_grid_20191231 %>%  
                                    st_transform(4326) %>%
                                    mutate( poverty_pct = 100 * poverty_rate , 
                                            unemployment_pct = 100 * unemployment_rate ,
                                                Label = paste0( "<br><b>Demographics :</b>", 
                                                   "<br>",
                                                   "<table>" ,
                                                   "<tr><td>", "Poverty Rate%:", "</td><td>"  , round(poverty_pct, 1), "</td></tr>" ,
                                                   "<tr><td>", "      Income:", "</td><td>$"  , round(income,0) , "</td></tr>" ,
                                                   "<tr><td>", "Unemployment%:", "</td><td>"  , round( unemployment_pct , 1) , "</td></tr>" ,
                                                   "<tr><td>", "      Density #/sq mile:", "</td><td>"  , round(density,0) , "</td></tr>" ,
                                                   "</table>"  )) 


cell_palette <- colorBin("Spectral", hotspot_cells_20191231$rank , 5 )

poverty_palette <- colorNumeric("Greens", demographics_cells_20191231$poverty_pct , 5 )

density_palette <- colorNumeric("Oranges", demographics_cells_20191231$density , 5 )

income_palette <- colorNumeric("RdYlBu", demographics_cells_20191231$income , 5 )



```


The Leaflet map below depicts Raleigh with multiple layers.  It focuses on show the Random Forest Hotspot predictions for the test period Jan 1, 2020 - Dec 31, 2020.   The observation date is Dec 31, 2019.
The map compares the following datasets.   Note that all demographic variables are calculated in similar fashion: from the ACS 2019 survey at the Census Block Group level and spatially interpolated to each grid cell.

*    `Hotspots Predicted`  Crime hotspots - defined as the top 100 grid cells ranked by predicted assaults as calculated by the model.  
*    `Poverty`  - expressed as a percentage of the population in the grid cell.
*    `Household Median Income` in 2019 inflated adjusted dollars. 
*    `Population Density` - expressed as number of people per square mile.


```{r}

RF_Hotspots_20191231 <- leaflet(hotspot_cells_20191231 , 
                                width = "100%"  , 
                                height = "1080px"
                                ) %>%
    setView( lng = -78.6382,  lat = 35.83 , zoom = 12) %>%
  addPolygons(color = "#444444", weight = 1, smoothFactor = 0.5,
    opacity = 1.0, fillOpacity = 0.75,
    fillColor = ~cell_palette(rank) ,
    highlightOptions = highlightOptions(color = "yellow", 
                                       weight = 3,
                                       bringToFront = TRUE), 
      label = ~lapply(Label, htmltools::HTML) ,
      labelOptions = labelOptions(noHide = FALSE, direction = "bottom", style = list("font-size" = "15px", "box-shadow" = "3px 3px rgba(0,0,0,0.25)", "font-family" = "Helvetica")) ,
      group = "Hotspots Predicted"
    ) %>% 
  addLegend( pal = cell_palette , values = ~rank,  title = "Hotspots Predicted"  ) %>%
  
  addPolygons(data = demographics_cells_20191231,
              color = "#444444", weight = 1, smoothFactor = 0.5,
    opacity = 1.0, fillOpacity = 0.5,
    stroke = FALSE ,
    fillColor = ~poverty_palette(poverty_pct) ,
    highlightOptions = highlightOptions(color = "yellow", 
                                       weight = 3,
                                       bringToFront = TRUE), 
      label = ~lapply(Label, htmltools::HTML) ,
      labelOptions = labelOptions(noHide = FALSE, direction = "bottom", style = list("font-size" = "15px", "box-shadow" = "3px 3px rgba(0,0,0,0.25)", "font-family" = "Helvetica")) ,
      group = "Poverty"
    ) %>% 
  addLegend( data = demographics_cells_20191231,  pal = poverty_palette , values = ~poverty_pct,  title = "Poverty %"  ) %>%
  
  addPolygons(data = demographics_cells_20191231,
              color = "#444444", weight = 1, smoothFactor = 0.5,
    opacity = 1.0, fillOpacity = 0.5,
    stroke = FALSE ,
    fillColor = ~income_palette(income) ,
    highlightOptions = highlightOptions(color = "yellow", 
                                       weight = 3,
                                       bringToFront = TRUE), 
      label = ~lapply(Label, htmltools::HTML) ,
      labelOptions = labelOptions(noHide = FALSE, direction = "bottom", style = list("font-size" = "15px", "box-shadow" = "3px 3px rgba(0,0,0,0.25)", "font-family" = "Helvetica")) ,
      group = "Household Median Income"
    ) %>% 
  addLegend( data = demographics_cells_20191231,  pal = income_palette , values = ~income,  title = "Household <br>Median Income"  ) %>%

  addPolygons(data = demographics_cells_20191231,
              color = "#444444", weight = 1, smoothFactor = 0.5,
    opacity = 1.0, fillOpacity = 0.5,
    stroke = FALSE ,
    fillColor = ~density_palette(density) ,
    highlightOptions = highlightOptions(color = "yellow", 
                                       weight = 3,
                                       bringToFront = TRUE), 
      label = ~lapply(Label, htmltools::HTML) ,
      labelOptions = labelOptions(noHide = FALSE, direction = "bottom", style = list("font-size" = "15px", "box-shadow" = "3px 3px rgba(0,0,0,0.25)", "font-family" = "Helvetica")) ,
      group = "Population Density"
    ) %>% 
  addLegend( data = demographics_cells_20191231,  pal = density_palette , values = ~density,  title = "Population Density<br> sq. mile"  ) %>%

  
  addProviderTiles(providers$OpenStreetMap, group = "OSM") %>%
  addProviderTiles(providers$CartoDB.Positron, group = "CartoDB Positron") %>%
  addPolylines(data=gj_raleigh_buffer %>% st_transform(4326), 
               color='green' , weight=4, group="Raleigh Limits") %>%
  addLayersControl(baseGroups = c("CartoDB Positron", "OSM"),
                   overlayGroups = c("Raleigh Limits","Hotspots Predicted", "Poverty", "Household Median Income", "Population Density"),
                   options = layersControlOptions(collapsed = FALSE)) %>%
  hideGroup("Poverty") %>%
  hideGroup("Household Median Income") %>%
  hideGroup("Population Density") %>%
  addControl(rr, position = "bottomleft")


saveWidget(RF_Hotspots_20191231 ,paste0(data_dir, "RF_Demographics_Hotspots_20191231.html"), selfcontained = TRUE)

```



# Mapping Crime Hotspots to POIs and All Assaults

```{r load-pois}

pois_raleigh_file = paste0(data_dir, "EXP_POIS_RALEIGH.geojson")
pois_a_raleigh_file = paste0(data_dir, "EXP_POIS_A_RALEIGH.geojson")

pois_raleigh = sf::st_read(pois_raleigh_file, quiet = TRUE ) %>% st_transform(4326)
pois_a_raleigh = sf::st_read(pois_a_raleigh_file, quiet = TRUE ) %>% st_transform(4326)

```



```{r load-potholes}

cityworks_potholes_file = paste0(data_dir, "Cityworks_Potholes.geojson" )

gj_potholes = sf::st_read(cityworks_potholes_file, quiet = TRUE ) %>% st_transform(4326)

# We exclude potholdes initiated after 2018 to avoid look-ahead bias
# ------------------------------------------------------------------
gj_potholes %>% filter( year(as_date(initiate_date)) <= 2018 ) %>% filter( !st_is_empty(.)) -> gj_potholes_historic

gj_potholes_historic  %>% mutate( Label = paste0(   "Pothole" , "<br>",
                                                    "<table>" ,
                                                   "<tr><td>", "Start date:", "</td><td>"  , initiate_date, "</td></tr>" ,
                                                   "<tr><td>", "Location:", "</td><td>"  , location , "</td></tr>" ,
                                                   "<tr><td>", "workorder_id:","</td><td>" , workorder_id , "</td></tr>" ,
                                                   "</table>" ) ) -> gj_potholes_historic

```



```{r filter-fclass}

area_fclass = c("park", "swimming_pool", "playground", "hotel", "fast_food", "school", "restaurant", "bank", "track", "car_wash" , "sports_centre", "car_dealership", "supermarket", "golf_course" , 
                "museum", "department_store", "university", "dentist", "pharmacy", "cafe" , "motel", "beauty_shop" , "bar" , "theatre", "cinema", "post_office" , "hospital", "nightclub", "courthouse" ,"prison" , 
                "theme_park", "stadium" )

point_fclass = c("restaurant", "fast_food", "supermarket", "bank", "cafe", "hairdresser", "beauty_shop", "school", "bar" , "pub", "convenience" , "car_dealership", "department_store" , "pharmacy" ,  "stadium" , 
                 "hotel", "gift_shop", "university", "motel", "playground", "bakery" , "nightclub", "theatre", "cinema" )


union_fclass = sort(union( area_fclass, point_fclass))

pois_raleigh %>% filter( fclass %in% point_fclass) -> pois_to_use

pois_a_raleigh %>% filter( fclass %in% area_fclass) -> pois_a_to_use

```


```{r}

union_poi_palette <- colorFactor(palette = "plasma", domain = union_fclass )


pois_to_use %>% mutate( Label = paste0(  "<table>" ,
                                                   "<tr><td>", "POI Type:", "</td><td>"  , fclass, "</td></tr>" ,
                                                   "<tr><td>", "Name:", "</td><td>"  , name , "</td></tr>" ,
                                                   "<tr><td>", "osm_id:","</td><td>" , osm_id , "</td></tr>" ,
                                                   "</table>" ) ) -> pois_to_use


pois_a_to_use %>% mutate( Label = paste0(  "<table>" ,
                                                   "<tr><td>", "POI Type:", "</td><td>"  , fclass, "</td></tr>" ,
                                                   "<tr><td>", "Name:", "</td><td>"  , name , "</td></tr>" ,
                                                   "<tr><td>", "osm_id:","</td><td>" , osm_id , "</td></tr>" ,
                                                   "</table>" ) ) -> pois_a_to_use


```


## Random Forests - POI

```{r}



rpoi <- tags$div(HTML('<h3>Assaults: Actual and Predicted</h3><br>2020 Forecasts + Realized<br>with Points of Interest, Potholes<br>Random Forest Model<br>Raleigh, NC<br>Alexander Ng' ) )


crime_palette <- colorBin("Reds",  bins = c( 0, 1, 3, 5, 10, 30, 40, 80) , na.color = "red" )


actual_crime_palette <- colorBin("Blues",  bins = c( 0, 1, 3, 5, 10, 30, 40, 80) , na.color = "blue" )

crime_labels <- c( "<= 1", "<= 3", "<= 5", "<=10", "<=30", "<=40", "<=80", ">80")

demographics_cells_20191231 %>%  mutate( Label = paste0( "<br><b>Hotspot :</b>",rank , "</b>", 
                                 
                                                   "<br><b>","Assaults 2020 ", "</b>" ,
                                                   "<table>" ,
                                                   "<tr><td>", "Predicted:", "</td><td>"    , round(pred_crime_freq, 1) , "</td></tr>" ,
                                                   "<tr><td>", "       2020:", "</td><td>"  , crime_freq , "</td></tr>" ,
                                                   "<tr><td>", "       2019:", "</td><td>"  , round(0.5 * ( crime_freq_6M + crime_freq_1Y), 0) , "</td></tr>" ,
                                                   "<tr><td>", "       2018:", "</td><td>"  , crime_freq_2Y , "</td></tr>" ,
                                                   "<tr><td>", "       2017:", "</td><td>"  , crime_freq_3Y, "</td></tr>" ,
                                                   "</table>"
                                                            )) -> demographics_cells_20191231


POI_Hotspots_20191231 <- leaflet(hotspot_cells_20191231 , 
                                width = "100%"  , 
                                height = "1080px"
                                ) %>%
    setView( lng = -78.6382,  lat = 35.83 , zoom = 12) %>%
  
    addPolygons(data = demographics_cells_20191231,
              color = "lightgray", weight = 1, smoothFactor = 0.5,
    opacity = 1.0, fillOpacity = 0.5,
    stroke = TRUE ,
    fillColor = ~crime_palette(pred_crime_freq) ,
    highlightOptions = highlightOptions(color = "yellow", 
                                       weight = 3,
                                       bringToFront = TRUE), 
      label = ~lapply(Label, htmltools::HTML) ,
      labelOptions = labelOptions(noHide = FALSE, direction = "bottom", style = list("font-size" = "15px", "box-shadow" = "3px 3px rgba(0,0,0,0.25)", "font-family" = "Helvetica")) ,
      group = "Assaults Predicted"
    ) %>% 
  addLegend( data = demographics_cells_20191231,  pal = crime_palette , values = ~pred_crime_freq ,  title = "All Assaults<br>Predicted"  , position = "topleft") %>%
    addPolygons(data = demographics_cells_20191231,
              color = "lightgray", weight = 1, smoothFactor = 0.5,
    opacity = 1.0, fillOpacity = 0.5,
    stroke = TRUE ,
    fillColor = ~actual_crime_palette(crime_freq) ,
    highlightOptions = highlightOptions(color = "yellow", 
                                       weight = 3,
                                       bringToFront = TRUE), 
      label = ~lapply(Label, htmltools::HTML) ,
      labelOptions = labelOptions(noHide = FALSE, direction = "bottom", style = list("font-size" = "15px", "box-shadow" = "3px 3px rgba(0,0,0,0.25)", "font-family" = "Helvetica")) ,
      group = "Assaults Actual"
    ) %>% 
  addLegend( data = demographics_cells_20191231,  pal = actual_crime_palette , values = ~crime_freq ,  title = "All Assaults<br>Actual" , position = "topleft" ) %>%
  addPolygons(data = hotspot_cells_20191231 , 
              color = "#444444", weight = 1, smoothFactor = 0.5,
    opacity = 1.0, fillOpacity = 0.75,
    fillColor = ~cell_palette(rank) ,
    highlightOptions = highlightOptions(color = "yellow", 
                                       weight = 3,
                                       bringToFront = TRUE), 
      label = ~lapply(Label, htmltools::HTML) ,
      labelOptions = labelOptions(noHide = FALSE, direction = "bottom", style = list("font-size" = "15px", "box-shadow" = "3px 3px rgba(0,0,0,0.25)", "font-family" = "Helvetica")) ,
      group = "Hotspots Predicted"
    ) %>% 
  addLegend( pal = cell_palette , values = ~rank,   title = "Top 100 Hotspots<br>Ranked"  ) %>%

  addPolygons( data = pois_a_to_use , 
               color = "#444444", weight = 1, smoothFactor = 0.5,
               opacity = 1.0, fillOpacity = 0.5,
               fillColor = "orange" ,
               highlightOptions = highlightOptions(color = "yellow", 
                                       weight = 3,
                                       bringToFront = TRUE), 
               
               labelOptions = labelOptions(noHide = FALSE, direction = "bottom", style = list("font-size" = "15px", "box-shadow" = "3px 3px rgba(0,0,0,0.25)", "font-family" = "Helvetica")) ,
               label = ~lapply(Label, htmltools::HTML), group = "Points of Interest - A") %>%
  
  addCircleMarkers( data = gj_potholes_historic, label = ~lapply(Label, htmltools::HTML), 
                    labelOptions = labelOptions(noHide = FALSE, direction = "bottom", style = list("font-size" = "15px", "box-shadow" = "3px 3px rgba(0,0,0,0.25)", "font-family" = "Helvetica")) ,
                    color = "red", radius = 4, group = "Potholes" ) %>%
  addCircleMarkers( data = pois_to_use, label = ~lapply(Label, htmltools::HTML), 
                    labelOptions = labelOptions(noHide = FALSE, direction = "bottom", style = list("font-size" = "15px", "box-shadow" = "3px 3px rgba(0,0,0,0.25)", "font-family" = "Helvetica")) ,
                    color = ~union_poi_palette(fclass), radius = 5, group = "Points of Interest - P" ) %>%
  addLegend( data  = pois_to_use, pal = union_poi_palette , values = ~fclass ,   title = "Points of Interest"  ) %>%
  
  
  addProviderTiles(providers$OpenStreetMap, group = "OSM") %>%
  addProviderTiles(providers$CartoDB.Positron, group = "CartoDB Positron") %>%
  addPolylines(data=gj_raleigh_buffer %>% st_transform(4326), 
               color='green' , weight=4, group="Raleigh Limits") %>%
  addLayersControl(baseGroups = c("CartoDB Positron", "OSM"),
                   overlayGroups = c("Raleigh Limits","Hotspots Predicted" , "Assaults Predicted", "Assaults Actual", "Points of Interest - P", "Points of Interest - A", "Potholes"),
                   options = layersControlOptions(collapsed = FALSE)) %>%
  hideGroup("Potholes") %>%
  hideGroup("Points of Interest - P") %>%
  hideGroup("Points of Interest - A") %>%
  hideGroup("Assaults Predicted") %>%
  hideGroup("Hotspots Predicted") %>%
  addControl(rpoi, position = "bottomleft")


saveWidget(POI_Hotspots_20191231 ,paste0(data_dir, "RF_POI_Hotspots_20191231.html" ), selfcontained = TRUE)


```



## Cubist Leaflet Map - Demographics

```{r load-cu-grid-output}

# Load the grid cells from random forest. Add HTML table as label when hovering.
# ---------------------------------------------------------------------------------
hotspot_cells_20191231 <- PAI_cu_ranked_grid_20191231 %>% 
                                    filter( hotspot==1) %>% 
                                    st_transform(4326) %>% 
                                    mutate( Label = paste0( "<br><b>Hotspot :</b>",rank , "</b>", 
                                                   "<br><b>","Assaults 2020 ", "</b>" ,
                                                   "<table>" ,
                                                   "<tr><td>", "Predicted:", "</td><td>"    , round(pred_crime_freq, 1) , "</td></tr>" ,
                                                   "<tr><td>", "       2020:", "</td><td>"  , crime_freq , "</td></tr>" ,
                                                   "<tr><td>", "       2019:", "</td><td>"  , round(0.5 * ( crime_freq_6M + crime_freq_1Y), 0) , "</td></tr>" ,
                                                   "<tr><td>", "       2018:", "</td><td>"  , crime_freq_2Y , "</td></tr>" ,
                                                   "<tr><td>", "       2017:", "</td><td>"  , crime_freq_3Y, "</td></tr>" ,
                                                   "</table>"
                                                            ))

#
# Save the crime grid.  Add HTML table for label when hovering over each cell.
# ---------------------------------------------------------------------------------
demographics_cells_20191231 <-  PAI_cu_ranked_grid_20191231 %>%  
                                    st_transform(4326) %>%
                                    mutate( poverty_pct = 100 * poverty_rate , 
                                            unemployment_pct = 100 * unemployment_rate ,
                                                Label = paste0( "<br><b>Demographics :</b>", 
                                                   "<br>",
                                                   "<table>" ,
                                                   "<tr><td>", "Poverty Rate%:", "</td><td>"  , round(poverty_pct, 1), "</td></tr>" ,
                                                   "<tr><td>", "      Income:", "</td><td>$"  , round(income,0) , "</td></tr>" ,
                                                   "<tr><td>", "Unemployment%:", "</td><td>"  , round( unemployment_pct , 1) , "</td></tr>" ,
                                                   "<tr><td>", "      Density #/sq mile:", "</td><td>"  , round(density,0) , "</td></tr>" ,
                                                   "</table>"  )) 


cell_palette <- colorBin("Spectral", hotspot_cells_20191231$rank , 5 )

poverty_palette <- colorNumeric("Greens", demographics_cells_20191231$poverty_pct , 5 )

density_palette <- colorNumeric("Oranges", demographics_cells_20191231$density , 5 )

income_palette <- colorNumeric("RdYlBu", demographics_cells_20191231$income , 5 )



```


```{r}

CU_Hotspots_20191231 <- leaflet(hotspot_cells_20191231 , 
                                width = "100%"  , 
                                height = "1080px"
                                ) %>%
    setView( lng = -78.6382,  lat = 35.83 , zoom = 12) %>%
  addPolygons(color = "#444444", weight = 1, smoothFactor = 0.5,
    opacity = 1.0, fillOpacity = 0.75,
    fillColor = ~cell_palette(rank) ,
    highlightOptions = highlightOptions(color = "yellow", 
                                       weight = 3,
                                       bringToFront = TRUE), 
      label = ~lapply(Label, htmltools::HTML) ,
      labelOptions = labelOptions(noHide = FALSE, direction = "bottom", style = list("font-size" = "15px", "box-shadow" = "3px 3px rgba(0,0,0,0.25)", "font-family" = "Helvetica")) ,
      group = "Hotspots Predicted"
    ) %>% 
  addLegend( pal = cell_palette , values = ~rank,  title = "Hotspots Predicted"  ) %>%
  
  addPolygons(data = demographics_cells_20191231,
              color = "#444444", weight = 1, smoothFactor = 0.5,
    opacity = 1.0, fillOpacity = 0.5,
    stroke = FALSE ,
    fillColor = ~poverty_palette(poverty_pct) ,
    highlightOptions = highlightOptions(color = "yellow", 
                                       weight = 3,
                                       bringToFront = TRUE), 
      label = ~lapply(Label, htmltools::HTML) ,
      labelOptions = labelOptions(noHide = FALSE, direction = "bottom", style = list("font-size" = "15px", "box-shadow" = "3px 3px rgba(0,0,0,0.25)", "font-family" = "Helvetica")) ,
      group = "Poverty"
    ) %>% 
  addLegend( data = demographics_cells_20191231,  pal = poverty_palette , values = ~poverty_pct,  title = "Poverty %"  ) %>%
  
  addPolygons(data = demographics_cells_20191231,
              color = "#444444", weight = 1, smoothFactor = 0.5,
    opacity = 1.0, fillOpacity = 0.5,
    stroke = FALSE ,
    fillColor = ~income_palette(income) ,
    highlightOptions = highlightOptions(color = "yellow", 
                                       weight = 3,
                                       bringToFront = TRUE), 
      label = ~lapply(Label, htmltools::HTML) ,
      labelOptions = labelOptions(noHide = FALSE, direction = "bottom", style = list("font-size" = "15px", "box-shadow" = "3px 3px rgba(0,0,0,0.25)", "font-family" = "Helvetica")) ,
      group = "Household Median Income"
    ) %>% 
  addLegend( data = demographics_cells_20191231,  pal = income_palette , values = ~income,  title = "Household <br>Median Income"  ) %>%

  addPolygons(data = demographics_cells_20191231,
              color = "#444444", weight = 1, smoothFactor = 0.5,
    opacity = 1.0, fillOpacity = 0.5,
    stroke = FALSE ,
    fillColor = ~density_palette(density) ,
    highlightOptions = highlightOptions(color = "yellow", 
                                       weight = 3,
                                       bringToFront = TRUE), 
      label = ~lapply(Label, htmltools::HTML) ,
      labelOptions = labelOptions(noHide = FALSE, direction = "bottom", style = list("font-size" = "15px", "box-shadow" = "3px 3px rgba(0,0,0,0.25)", "font-family" = "Helvetica")) ,
      group = "Population Density"
    ) %>% 
  addLegend( data = demographics_cells_20191231,  pal = density_palette , values = ~density,  title = "Population Density<br> sq. mile"  ) %>%

  
  addProviderTiles(providers$OpenStreetMap, group = "OSM") %>%
  addProviderTiles(providers$CartoDB.Positron, group = "CartoDB Positron") %>%
  addPolylines(data=gj_raleigh_buffer %>% st_transform(4326), 
               color='green' , weight=4, group="Raleigh Limits") %>%
  addLayersControl(baseGroups = c("CartoDB Positron", "OSM"),
                   overlayGroups = c("Raleigh Limits","Hotspots Predicted", "Poverty", "Household Median Income", "Population Density"),
                   options = layersControlOptions(collapsed = FALSE)) %>%
  hideGroup("Poverty") %>%
  hideGroup("Household Median Income") %>%
  hideGroup("Population Density") %>%
  addControl(rc, position = "bottomleft")


saveWidget(CU_Hotspots_20191231 ,paste0(data_dir, "CU_Demographics_Hotspots_20191231.html"), selfcontained = TRUE)

```




## Cubist - POI

```{r}



cpoi <- tags$div(HTML('<h3>Assaults: Actual and Predicted</h3><br>2020 Forecasts + Realized<br>with Points of Interest, Potholes<br>Cubist Model<br>Raleigh, NC<br>Alexander Ng' ) )


crime_palette <- colorBin("Reds",  bins = c( 0, 1, 3, 5, 10, 30, 40, 80) , na.color = "red" )


actual_crime_palette <- colorBin("Blues",  bins = c( 0, 1, 3, 5, 10, 30, 40, 80) , na.color = "blue" )

crime_labels <- c( "<= 1", "<= 3", "<= 5", "<=10", "<=30", "<=40", "<=80", ">80")

demographics_cells_20191231 %>%  mutate( Label = paste0( "<br><b>Hotspot :</b>",rank , "</b>", 
                                 
                                                   "<br><b>","Assaults 2020 ", "</b>" ,
                                                   "<table>" ,
                                                   "<tr><td>", "Predicted:", "</td><td>"    , round(pred_crime_freq, 1) , "</td></tr>" ,
                                                   "<tr><td>", "       2020:", "</td><td>"  , crime_freq , "</td></tr>" ,
                                                   "<tr><td>", "       2019:", "</td><td>"  , round(0.5 * ( crime_freq_6M + crime_freq_1Y), 0) , "</td></tr>" ,
                                                   "<tr><td>", "       2018:", "</td><td>"  , crime_freq_2Y , "</td></tr>" ,
                                                   "<tr><td>", "       2017:", "</td><td>"  , crime_freq_3Y, "</td></tr>" ,
                                                   "</table>"
                                                            )) -> demographics_cells_20191231


POI_Hotspots_20191231 <- leaflet(hotspot_cells_20191231 , 
                                width = "100%"  , 
                                height = "1080px"
                                ) %>%
    setView( lng = -78.6382,  lat = 35.83 , zoom = 12) %>%
  
    addPolygons(data = demographics_cells_20191231,
              color = "lightgray", weight = 1, smoothFactor = 0.5,
    opacity = 1.0, fillOpacity = 0.5,
    stroke = TRUE ,
    fillColor = ~crime_palette(pred_crime_freq) ,
    highlightOptions = highlightOptions(color = "yellow", 
                                       weight = 3,
                                       bringToFront = TRUE), 
      label = ~lapply(Label, htmltools::HTML) ,
      labelOptions = labelOptions(noHide = FALSE, direction = "bottom", style = list("font-size" = "15px", "box-shadow" = "3px 3px rgba(0,0,0,0.25)", "font-family" = "Helvetica")) ,
      group = "Assaults Predicted"
    ) %>% 
  addLegend( data = demographics_cells_20191231,  pal = crime_palette , values = ~pred_crime_freq ,  title = "All Assaults<br>Predicted"  , position = "topleft") %>%
    addPolygons(data = demographics_cells_20191231,
              color = "lightgray", weight = 1, smoothFactor = 0.5,
    opacity = 1.0, fillOpacity = 0.5,
    stroke = TRUE ,
    fillColor = ~actual_crime_palette(crime_freq) ,
    highlightOptions = highlightOptions(color = "yellow", 
                                       weight = 3,
                                       bringToFront = TRUE), 
      label = ~lapply(Label, htmltools::HTML) ,
      labelOptions = labelOptions(noHide = FALSE, direction = "bottom", style = list("font-size" = "15px", "box-shadow" = "3px 3px rgba(0,0,0,0.25)", "font-family" = "Helvetica")) ,
      group = "Assaults Actual"
    ) %>% 
  addLegend( data = demographics_cells_20191231,  pal = actual_crime_palette , values = ~crime_freq ,  title = "All Assaults<br>Actual" , position = "topleft" ) %>%
  addPolygons(data = hotspot_cells_20191231 , 
              color = "#444444", weight = 1, smoothFactor = 0.5,
    opacity = 1.0, fillOpacity = 0.75,
    fillColor = ~cell_palette(rank) ,
    highlightOptions = highlightOptions(color = "yellow", 
                                       weight = 3,
                                       bringToFront = TRUE), 
      label = ~lapply(Label, htmltools::HTML) ,
      labelOptions = labelOptions(noHide = FALSE, direction = "bottom", style = list("font-size" = "15px", "box-shadow" = "3px 3px rgba(0,0,0,0.25)", "font-family" = "Helvetica")) ,
      group = "Hotspots Predicted"
    ) %>% 
  addLegend( pal = cell_palette , values = ~rank,   title = "Top 100 Hotspots<br>Ranked"  ) %>%

  addPolygons( data = pois_a_to_use , 
               color = "#444444", weight = 1, smoothFactor = 0.5,
               opacity = 1.0, fillOpacity = 0.5,
               fillColor = "orange" ,
               highlightOptions = highlightOptions(color = "yellow", 
                                       weight = 3,
                                       bringToFront = TRUE), 
               
               labelOptions = labelOptions(noHide = FALSE, direction = "bottom", style = list("font-size" = "15px", "box-shadow" = "3px 3px rgba(0,0,0,0.25)", "font-family" = "Helvetica")) ,
               label = ~lapply(Label, htmltools::HTML), group = "Points of Interest - A") %>%
  
  addCircleMarkers( data = gj_potholes_historic, label = ~lapply(Label, htmltools::HTML), 
                    labelOptions = labelOptions(noHide = FALSE, direction = "bottom", style = list("font-size" = "15px", "box-shadow" = "3px 3px rgba(0,0,0,0.25)", "font-family" = "Helvetica")) ,
                    color = "red", radius = 4, group = "Potholes" ) %>%
  addCircleMarkers( data = pois_to_use, label = ~lapply(Label, htmltools::HTML), 
                    labelOptions = labelOptions(noHide = FALSE, direction = "bottom", style = list("font-size" = "15px", "box-shadow" = "3px 3px rgba(0,0,0,0.25)", "font-family" = "Helvetica")) ,
                    color = ~union_poi_palette(fclass), radius = 5, group = "Points of Interest - P" ) %>%
  addLegend( data  = pois_to_use, pal = union_poi_palette , values = ~fclass ,   title = "Points of Interest"  ) %>%
  
  
  addProviderTiles(providers$OpenStreetMap, group = "OSM") %>%
  addProviderTiles(providers$CartoDB.Positron, group = "CartoDB Positron") %>%
  addPolylines(data=gj_raleigh_buffer %>% st_transform(4326), 
               color='green' , weight=4, group="Raleigh Limits") %>%
  addLayersControl(baseGroups = c("CartoDB Positron", "OSM"),
                   overlayGroups = c("Raleigh Limits","Hotspots Predicted" , "Assaults Predicted", "Assaults Actual", "Points of Interest - P", "Points of Interest - A", "Potholes"),
                   options = layersControlOptions(collapsed = FALSE)) %>%
  hideGroup("Potholes") %>%
  hideGroup("Points of Interest - P") %>%
  hideGroup("Points of Interest - A") %>%
  hideGroup("Assaults Predicted") %>%
  hideGroup("Hotspots Predicted") %>%
  addControl(cpoi, position = "bottomleft")


saveWidget(POI_Hotspots_20191231 ,paste0(data_dir, "CU_POI_Hotspots_20191231.html" ), selfcontained = TRUE)


```

## TAB 2

### Cubist POI Hotspots 2020

```{r out.width='100%'}

POI_Hotspots_20191231
```
## TAB 3

### Cubist Demographics 2020

```{r out.width = '100%'}
CU_Hotspots_20191231

```
