---
title: "PART 1: EDA of Census Data and POI"
author: "Alexander Ng"
date: "11/25/2021"
output: 
  html_document:
    df_print: paged
    highlight: tango
    number_sections: yes
    theme: readable
    toc: yes
    toc_depth: 2
    toc_float: no
    fontsize: 12

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction

This module extracts and describes the Census data and Points of Interest to be used for Model Evaluation.

The raw data files were downloaded in geojson format from the American Community Survey 2019 - 5 year survey.
The ultimate data source is the US Census Bureau.  The data distributor is censusreporter.org


```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(kableExtra)
library(sf)
library(stars)
library(ggplot2)
library(cowplot)
library(ggspatial)
library(viridis)
library(leaflet)
library(classInt)
library(RColorBrewer)
```

```{r echo=FALSE}
root_dir = "/Volumes/GDRIVE_SSD/homes/alex/datascience/698_CAPSTONE_2021_FALL/"

data_dir = paste0(root_dir , "project_data/")

working_dir = paste0(root_dir, "project_code/Part1_EDA/")

```


# City Limits

Let's load the Raleigh Corporate Limits.   By superimposing the city limits with census data, we can verify that the city is fully covered by the latter.


```{r map-file-load}
raleigh_corporate_limits_file = paste0(data_dir, "Corporate_Limits.geojson")


# Load the geojson files
# change the coordinate system to 2264
# filter objects that belong to RALEIGHT
# --------------------------------------------
gj_raleigh <- sf::st_read(raleigh_corporate_limits_file, quiet = TRUE) %>% 
  st_transform(2264) %>% 
  filter( SHORT_NAME == 'RAL' )

un_raleigh_buffer = st_buffer(gj_raleigh, dist = 100 ) %>% st_union() %>% st_sf()

```


# Poverty Status (B17017)

Load the poverty geojson census data.

```{r}
pov_root = "acs2019_5yr_B17017_15000US371830529022"

poverty_B17017_raw = st_read( paste0(data_dir, pov_root, "/", pov_root, ".geojson" )
                                     , quiet = TRUE ) %>% st_transform(2264)
```

Extract the regional average poverty rate.

```{r}
regional_avg_poverty_rate = poverty_B17017_raw %>% 
  filter( geoid == "31000US39580") %>% 
  mutate( rate = B17017002 / B17017001)  %>% 
  pull(rate)

```

Impute any missing poverty rate regions with the median rate for the entire Raleigh-Cary Metro Area.

```{r}

poverty_B17017_raw %>% 
  select( geoid, name, B17017001, B17017002 ) %>% 
  filter( name  != 'Raleigh-Cary, NC Metro Area')  %>% 
  mutate( poverty_rate = ifelse( B17017001 == 0, regional_avg_poverty_rate , B17017002 / B17017001) )  %>% 
  mutate( area = st_area(.)) -> gj_poverty

```




```{r}

pal_fun = colorQuantile("Spectral", NULL , n = 6)

gj_poverty_leaf = st_transform(gj_poverty, 4326)

p_popup <- paste0("Poverty: ", format(round(gj_poverty$poverty_rate, 3), nsmall = 3 ), "<br>" ,
                  "Name: ", gj_poverty$name , "<br>" ,
                  "Geoid: ", gj_poverty$geoid, "<br>" )
                  

# get quantile breaks. Add .00001 offset to catch the lowest value
breaks_qt <- classIntervals(c(min(gj_poverty$poverty_rate) - .00001, 
                              gj_poverty$poverty_rate), n = 6, style = "quantile")

breaks_qt
```

```{r}

leaflet( gj_poverty_leaf ) %>% addPolygons( stroke = FALSE, # Remove polygon borders
                                                 fillColor = ~pal_fun(poverty_rate),  # set fill color with function from above and value
                                                 fillOpacity = 0.5 ,
                                                 smoothFactor = 0.5, # make it nicer
                                                 popup = p_popup
                                                 ) %>% addTiles() %>%
       addLegend( "bottomright", # Location
                    colors = brewer.pal(6, "Spectral") ,
                  labels = paste0("up to ", format(breaks_qt$brks[-1], digits = 2)) ,
                  title = "Raleigh Poverty rate by Census Block Group"
                  )
 

```


# Unemployment (B23025)

Table `B23025` contains Employment Status for the Population 16 years and over.   The relevant field is `Unemployed %` from the civilian labor force.


```{r}

unemp_root = "acs2019_5yr_B23025_15000US371830529022"

unemp_B23025_raw = st_read( paste0(data_dir, unemp_root, "/", unemp_root, ".geojson" )
                                     , quiet = TRUE ) %>% st_transform(2264)
```


Extract the regional average unemployment rate for imputation of regions with undefined unemployment rates.

```{r}
regional_avg_unemp_rate  = unemp_B23025_raw %>% 
  filter( geoid == "31000US39580") %>% 
  mutate( rate = B23025005 / B23025001)  %>% 
  pull( rate )

```

Impute any missing poverty rate regions with the median rate for the entire Raleigh-Cary Metro Area.

```{r}

unemp_B23025_raw %>% 
  select( geoid, name, B23025005, B23025001 ) %>% 
  filter( name  != 'Raleigh-Cary, NC Metro Area')  %>% 
  mutate( unemployment_rate = ifelse( B23025001  == 0, regional_avg_unemp_rate , B23025005 / B23025001) )  %>% 
  mutate( area = st_area(.)) -> gj_unemp

```


```{r}

gj_unemp_leaf = st_transform(gj_unemp, 4326)

p_popup <- paste0("Unemp: ", format(round(100 * gj_unemp$unemployment_rate, 1), nsmall = 1 ), "<br>" ,
                  "Name: ", gj_unemp$name , "<br>" ,
                  "Geoid: ", gj_unemp$geoid, "<br>" )
                  

# get quantile breaks. Add .00001 offset to catch the lowest value
breaks_qt <- classIntervals(c(min(gj_unemp$unemployment_rate) - .00001, 
                              gj_unemp$unemployment_rate), n = 6, style = "quantile")

breaks_qt
```


```{r}

leaflet( gj_unemp_leaf ) %>% addPolygons( stroke = FALSE, # Remove polygon borders
                                                 fillColor = ~pal_fun(unemployment_rate),  # set fill color with function from above and value
                                                 fillOpacity = 0.5 ,
                                                 smoothFactor = 0.5, # make it nicer
                                                 popup = p_popup
                                                 ) %>% addTiles() %>%
       addLegend( "bottomright", # Location
                    colors = brewer.pal(6, "Spectral") ,
                  labels = paste0("up to ", format(breaks_qt$brks[-1], digits = 2)) ,
                  title = "Raleigh Unemployment By Block Group"
                  )
 

```



```{r eval = FALSE}
ggplot() + geom_sf(data=gj_unemp , aes( fill = unemployment_rate  )  ) +
  scale_fill_viridis_c( alpha = .4) +
  annotation_map_tile(type="osm", progress = "none", alpha = 0.5 ) 

```

# Household Income (B19013)


Table `B19013` contains Median Household Income .   The relevant field is `Unemployed %` from the civilian labor force.


```{r}

#income_root = "acs2019_5yr_B19013_15000US371830523011"   # Smaller Income region

income_root = "acs2019_5yr_B19013_15000US371830529022"   # Larger Income region


income_B19013_raw = st_read( paste0(data_dir, income_root, "/", income_root, ".geojson" )
                                     , quiet = TRUE ) %>% st_transform(2264)
```


Extract the regional average unemployment rate for imputation of regions with undefined unemployment rates.

```{r}
regional_avg_income  = 67266  # Median household income for Raleigh city in 2019 inflated adjusted dollars

```


Impute any missing median household income with the regional median household income for the entire Raleigh-Cary Metro Area.

```{r}

income_B19013_raw %>%  select( geoid, name, B19013001 ) %>% 
# filter( str_length(geoid) == 19 ) %>%  # Filter out the higher level aggregate regions shown by shorter geoid like Cary, NC, Raleigh, NC, etc.
 mutate( income = ifelse( is.na(B19013001) | B19013001  == 0, regional_avg_income , B19013001) )  %>% 
 mutate( area = st_area(.)) -> gj_income

```


```{r}

gj_income_leaf = st_transform(gj_income, 4326)

p_popup <- paste0("Income: $", format(round( gj_income$income, 0), nsmall = 0 ), "<br>" ,
                  "Name: ", gj_income$name , "<br>" ,
                  "Geoid: ", gj_income$geoid, "<br>" )
                  

# get quantile breaks. Add .00001 offset to catch the lowest value
breaks_qt <- classIntervals(c(min(gj_income$income) - .00001, 
                              gj_income$income), n = 6, style = "quantile")

breaks_qt
```



```{r}

leaflet( gj_income_leaf ) %>% addPolygons( stroke = FALSE, # Remove polygon borders
                                                 fillColor = ~pal_fun(income),  # set fill color with function from above and value
                                                 fillOpacity = 0.5 ,
                                                 smoothFactor = 0.5, # make it nicer
                                                 popup = p_popup
                                                 ) %>% addTiles() %>%
       addLegend( "bottomright", # Location
                    colors = brewer.pal(6, "Spectral") ,
                  labels = paste0("up to ", format(breaks_qt$brks[-1], digits = 2)) ,
                  title = "Raleigh Median Household Income"
                  )
 

```


# Population Density


Table `B01003` contains Population.   The relevant field is `B01003001` which is the total population of the census block group or region.


```{r}

pop_root = "acs2019_5yr_B01003_15000US371830529022"   # Larger Population region


pop_B01003_raw = st_read( paste0(data_dir, pop_root, "/", pop_root, ".geojson" )
                                     , quiet = TRUE ) %>% st_transform(2264)
```


There are no missing values in the Total Population table for any block group or region, so we will be able to define a population density as total population per square mile.

```{r}

pop_B01003_raw %>%  select( geoid, name, B01003001 ) %>% 
 mutate( population = ifelse( is.na(B01003001) , 0 , B01003001) )  %>% 
 mutate( area = st_area(.)) %>% 
 mutate( density = population / ( area / (5280^2 ) ) )    -> gj_pop

```



```{r}

gj_pop_leaf = st_transform(gj_pop, 4326)

p_popup <- paste0("Density: $", format(round( gj_pop$density, 0), nsmall = 0 ), "<br>" ,
                  "Name: ", gj_pop$name , "<br>" ,
                  "Geoid: ", gj_pop$geoid, "<br>" )
                  
```


# Checking Containment and Validity

First, we do a visual inspection using leaflet to ensure Raleigh is inside the census region.
Then, we'll run formal checks using `st_contains` for containment of the city in each census region
and `st_crosses` to confirm census regions have no overlapping areas.

The plot below examines Raleigh and median household income region.

```{r}

leaflet( gj_income_leaf ) %>% 
  addPolygons( data = gj_raleigh %>% st_transform( 4326 ) , weight = 2, color = "red", opacity = 0.8) %>%
  addPolygons( stroke = FALSE, # Remove polygon borders
                                                 fillColor = ~pal_fun(income),  # set fill color with function from above and value
                                                 fillOpacity = 0.5 ,
                                                 smoothFactor = 0.5, # make it nicer
                                                 popup = p_popup
                                                 )  %>%
                              addTiles() %>%
       addLegend( "bottomright", # Location
                    colors = brewer.pal(6, "Spectral") ,
                  labels = paste0("up to ", format(breaks_qt$brks[-1], digits = 2)) ,
                  title = "Raleigh Median Household Income"
                  )
 

```

We conclude Raleigh region is visually contained in the income region.

To do the next checks, we form a single multipolygon from the city `sf` object using `st_union` and likewise for the income data.


```{r}

un_raleigh = st_union(gj_raleigh)
un_income = st_union( gj_income)
```

Check the city has no area outside the income census region.  Area should be zero or empty.

```{r}
# st_difference is asymmetric.  We check if the city has any area outside the census attribute partition.
# If there is no area, st_area should not compute a positive area.
st_area( st_difference( un_raleigh, un_income))
```

Another check the income region contains the city.  
```{r}
# another check that census attribution partition wholly contains the city region
lengths( st_contains( un_income, un_raleigh ) )  > 0 
```

Next, check the census region is a proper 2-dimensional partition with no overlapping interiors.
This ensures that grid cells have well-defined census attribute based on 1 or more block group areas.
Answer should be `FALSE`.

```{r}
# The next line checks if any polygons in the sf table of the census attribute partition
# have overlapping 2-dimensional interiors. (st_crosses)
# lengths() - returns a list of length 0 or more corresponding to indices of polygons where intersection occurs
# lengths( x ) > 0 return a list checking if any list is length > 0 iff st_crosses returns no interior intersections
# any() scans the entire list of TRUE/FALSE for TRUE

any( lengths(st_crosses( gj_income) ) > 0  )  

```

Let's check poverty.  

```{r}

un_poverty = st_union(gj_poverty)

st_area( st_difference(un_raleigh, un_poverty))

lengths(st_contains(un_poverty, un_raleigh) ) > 0 


```
We conclude Raleigh region is entirely contained in the poverty region.

Let's confirm no overlap in the poverty `sf` region. Answer should be `FALSE`.

```{r}
any( lengths(st_crosses( gj_poverty) ) > 0  )
```


Let's check Raleigh is contained inside the unemployment data region.

```{r}
un_unemployment = st_union( gj_unemp)

st_area( st_difference(un_raleigh, un_unemployment))

lengths( st_contains(un_unemployment , un_raleigh) ) > 0 



```

We confirm that.  Finally, let's confirm no overlap in the `sf` unemployment region below.

```{r}
any( lengths(st_crosses( gj_unemp) ) > 0  )
```







# POI DATA

```{r}

poi_root = "north-carolina-latest-free.shp"


pois_free_raw = st_read( paste0(data_dir, poi_root, "/", "gis_osm_pois_free_1", ".shp" )
                                     , quiet = TRUE ) %>% st_transform(2264)


pois_a_free_raw = st_read( paste0(data_dir, poi_root, "/", "gis_osm_pois_a_free_1", ".shp" )
                                     , quiet = TRUE ) %>% st_transform(2264)

```

Use the bounding box of Raleigh to reduce the size of the POI dataset.

```{r}

bbox_raleigh_buffer = st_bbox(un_raleigh_buffer) %>% st_as_sfc() %>% st_transform(2264) # make a simple rectangle.

pois_free_raw[ bbox_raleigh_buffer, ] -> bb_pois_raleigh

pois_a_free_raw[ bbox_raleigh_buffer, ] -> bb_pois_a_raleigh
```

Now, use spatial intersection of the POI data with the Raleigh city buffered boundary to obtain the POIs within the city.

```{r}

pois_raleigh = bb_pois_raleigh %>% st_intersection(un_raleigh_buffer)

pois_a_raleigh = bb_pois_a_raleigh %>% st_intersection(un_raleigh_buffer)

```

```{r eval=TRUE, message = FALSE, error = FALSE}

# Note that sf guesses the format of the output file from the suffix.

st_write(pois_raleigh ,  dsn = paste0(data_dir, "EXP_POIS_RALEIGH.geojson"), delete_dsn = TRUE)

st_write(pois_a_raleigh ,  dsn = paste0(data_dir, "EXP_POIS_A_RALEIGH.geojson"), delete_dsn = TRUE)

```

```{r out.width="100%"}

pois_free_bbox_leaf = bb_pois_raleigh %>% st_transform(4326)

coords = st_coordinates(pois_free_bbox_leaf)

lat = coords[,2]
long = coords[,1]

leaflet( pois_free_bbox_leaf) %>% 
  addPolygons( data = gj_raleigh %>% st_transform( 4326 ) , weight = 2, color = "red", opacity = 0.5) %>%
                              addTiles() %>%
   addMarkers( data = pois_free_bbox_leaf,  popup = pois_free_bbox_leaf$fclass,  lng = long , lat = lat )


```

Let's evaluate the frequency table of the most common POIs for point objects.

```{r}
pois_raleigh %>% group_by( fclass) %>% summarize( ct = n()) %>% st_drop_geometry() %>% arrange( desc(ct)) %>%
  top_n( n = 30 )  %>%  ggplot( aes( x = reorder(fclass, ct),  y = ct)) + 
  geom_bar(stat = "identity") + coord_flip() + labs( x = "fclass", y = "Count", title = "Points of Interest in Raleigh" ) -> p1
  

pois_a_raleigh %>% group_by( fclass) %>% summarize( ct = n()) %>% st_drop_geometry() %>% arrange( desc(ct)) %>%
  top_n( n = 30 )  %>%  ggplot( aes( x = reorder(fclass, ct),  y = ct)) + 
  geom_bar(stat = "identity") + coord_flip() + labs( x = "fclass", y = "Count", title = "Areas of Interest in Raleigh" ) -> p2
  
plot_grid(p1, p2)

```

```{r}
pois_a_raleigh %>% filter( fclass == 'pitch') %>% mutate( area = st_area(.)) %>% st_drop_geometry() %>% View()

```
